---
title: "FA3"
author: "Aldwin Benidict Montilla"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 5: Sensor Relevance Analysis

A geospatial analysis system has four sensors supplying images. The percentage of images supplied by each sensor and the percentage of images relevant to a query are shown in the following table.

| Sensor | Percentage of Images Supplied | Percentage of Relevant Images |
|:------:|:----------------------------:|:-----------------------------:|
|   1    |              15              |              50               |
|   2    |              20              |              60               |
|   3    |              25              |              80               |
|   4    |              40              |              85               |

What is the overall percentage of relevant images?

### R Code & Solution

```{r sensor_analysis}
# Input the data into data vectors
sensor <- c(1, 2, 3, 4)
pct_supplied <- c(15, 20, 25, 40)
pct_relevant <- c(50, 60, 80, 85)

# Create a clean data frame to hold the info
sensor_data <- data.frame(Sensor = sensor, 
                          Pct_Supplied = pct_supplied, 
                          Pct_Relevant = pct_relevant)

# Convert percentages to proportions for probability calculation
p_sensor <- sensor_data$Pct_Supplied / 100
p_relevant_given_sensor <- sensor_data$Pct_Relevant / 100

# Calculate the overall probability using the Law of Total Probability
# P(Relevant) = Sum of [P(Sensor_i) * P(Relevant | Sensor_i)]
overall_prob_relevant <- sum(p_sensor * p_relevant_given_sensor)
overall_prob_relevant

# Convert back to a clean percentage format
overall_pct_relevant <- overall_prob_relevant * 100
```

```{r print}
# Print the result
print(paste0("The overall percentage of relevant images is: ", overall_pct_relevant, "%"))
```

### Interpretation
The total percentage of relevant images across the entire network of four sensors evaluates to **`r overall_pct_relevant` %**.



## Question 6: Probability Independence Verification

A fair coin is tossed twice. Let the sample space be $S = \{HH, HT, TH, TT\}$, where each outcome is equally likely. Consider the following three events:

*   **$E_1$**: The outcomes of the two tosses match ($HH$ or $TT$).
*   **$E_2$**: The first toss results in Heads ($HH$ or $HT$).
*   **$E_3$**: The second toss results in Heads ($TH$ or $HH$).

Show whether these three events are **pairwise independent** and whether they are **mutually independent**.


```{r independence_check}
# Define the complete sample space (S)
S <- c("HH", "HT", "TH", "TT")

# Define the each events (E)
E1 <- c("HH", "TT")
E2 <- c("HH", "HT")
E3 <- c("TH", "HH")

# Create a custom function to compute probabilities
get_prob <- function(event_set) {
  length(event_set) / length(S)
}
```

```{r prob_of_events}
# Compute individual probabilities
p_E1 <- get_prob(E1)
p_E2 <- get_prob(E2)
p_E3 <- get_prob(E3)

# Print individual probabilities clearly
print(paste("Probability of E1:", p_E1))
print(paste("Probability of E2:", p_E2))
print(paste("Probability of E3:", p_E3))
```

```{r pair_I}
# Check Pairwise Independence (Intersections)
int_1_2 <- intersect(E1, E2)
int_1_3 <- intersect(E1, E3)
int_2_3 <- intersect(E2, E3)

print(paste("Intersection of E1 and E2:", int_1_2))
print(paste("Intersection of E1 and E3:", int_1_3))
print(paste("Intersection of E2 and E3:", int_2_3))

```

```{r pairwise_I}

# Check if P(A and B) == P(A) * P(B)
p_int_1_2 <- get_prob(int_1_2) == (p_E1 * p_E2)
p_int_1_3 <- get_prob(int_1_3) == (p_E1 * p_E3)
p_int_2_3 <- get_prob(int_2_3) == (p_E2 * p_E3)


print(paste("Pair 1 & 2:", get_prob(int_1_2), "==", (p_E1 * p_E2)))
print(paste("Pair 1 & 3:", get_prob(int_1_3), "==", (p_E1 * p_E3)))
print(paste("Pair 2 & 3:", get_prob(int_2_3), "==", (p_E2 * p_E3)))

```

```{r mutual_I}
# Check Mutual Independence (Intersection of all three)
int_all <- intersect(intersect(E1, E2), E3)
p_int_all <- get_prob(int_all)
p_product_all <- p_E1 * p_E2 * p_E3
is_mutually_independent <- p_int_all == p_product_all


# Print results
print(paste("E1 & E2 independent?:", p_int_1_2))
print(paste("E1 & E3 independent?:", p_int_1_3))
print(paste("E2 & E3 independent?:", p_int_2_3))
print(paste("Mutually independent?:", is_mutually_independent))
```

### Conclusion

The simulation and mathematical analysis demonstrate that events $E_1$, $E_2$, and $E_3$ are **pairwise independent but not mutually independent**. 

When tested two at a time, knowing the outcome of one event leaves the probability of the other event completely unchanged at $50\%$. This satisfies the pairwise multiplication rule where $P(A \cap B) = P(A) \times P(B) = 0.25$.

However, they fail the test for mutual independence because the events share a strict logical relationship. If we know that the first coin is Heads ($E_2$) and the second coin is Heads ($E_3$), the probability that the two coins match ($E_1$) instantly jumps from $50\%$ to $100\%$. Because combining our knowledge of two events completely eliminates the randomness of the third event, the group multiplication rule breaks ($P(E_1 \cap E_2 \cap E_3) = 0.25$, which does not equal $P(E_1) \times P(E_2) \times P(E_3) = 0.125$).
